---
title: |
    | Summary Report of DecisionTree on bank_client_data
author: | 
    |     Achal, Gupta ---------------A00258772
date: "April 5, 2019"
output: html_document
editor_options: 
  chunk_output_type: inline
---
#Data Set Information:
The dataset is about direct marketing campaigns of a Portuguese banking institution. The method of data being collected in this marketing campaigns were based on phone calls. Often, more than one contact to the same client was required, in order to access if the product (bank term deposit) would be subscribed. 

#Attribute Information
###Input variables: bank client data:<br>
1 - **age: ** (numeric) <br>
2 - **job: **type of job (categorical: 'admin.','blue-collar','entrepreneur','housemaid','management','retired','self-employed','services','student','technician','unemployed','unknown')<br>
3 - **marital: ** marital status (categorical: 'divorced','married','single','unknown'; note: 'divorced' means divorced or widowed)<br>
4 - **education: ** (categorical: 'basic.4y','basic.6y','basic.9y','high.school','illiterate','professional.course','university.degree','unknown')<br>
5 - **default: **has credit in default? (categorical: 'no','yes','unknown')<br>
6 - **housing: **has housing loan? (categorical: 'no','yes','unknown')<br>
7 - **loan: **has personal loan? (categorical: 'no','yes','unknown')<br>

###related with the last contact of the current campaign:<br>
8 - **contact: **contact communication type (categorical: 'cellular','telephone') <br>
9 - **month: **last contact month of year (categorical: 'jan', 'feb', 'mar', ..., 'nov', 'dec')<br>
10 - **day_of_week: **last contact day of the week (categorical: 'mon','tue','wed','thu','fri')<br>
11 - **duration: **last contact duration, in seconds (numeric). Important note: this attribute highly affects the output target (e.g., if duration=0 then y='no'). Yet, the duration is not known before a call is performed. Also, after the end of the call y is obviously known. Thus, this input should only be included for benchmark purposes and should be discarded if the intention is to have a realistic predictive model.<br>

###other attributes:<br>
12 - **campaign: **number of contacts performed during this campaign and for this client (numeric, includes last contact)<br>
13 - **pdays: **number of days that passed by after the client was last contacted from a previous campaign (numeric; 999 means client was not previously contacted)<br>
14 - **previous: **number of contacts performed before this campaign and for this client (numeric)<br>
15 - **poutcome: **outcome of the previous marketing campaign (categorical: 'failure','nonexistent','success')<br>

###social and economic context attributes: <br>
16 - **emp.var.rate: **employment variation rate - quarterly indicator (numeric)<br>
17 - **cons.price.idx: **consumer price index - monthly indicator (numeric) <br>
18 - **cons.conf.idx: **consumer confidence index - monthly indicator (numeric) <br>
19 - **euribor3m: **euribor 3 month rate - daily indicator (numeric)<br>
20 - **nr.employed: **number of employees - quarterly indicator (numeric)<br>

###Output variable (desired target):<br>
21 - **y** - has the client subscribed a term deposit? (binary: 'yes','no')<br>
<br>
<br>

```{r packagesLibrary, include=FALSE, warning = FALSE, message = FALSE}
###IMPORTING LIBRARIES
#Below code will import all the required libraries.

#install.packages("ggplot2")
#install.packages("rpart")
#install.packages("rattle")
#install.packages("caret")
#install.packages("corrplot")
library(ggplot2)
library(rpart)
library(rpart.plot)
library(rattle)
library(caret)
library(corrplot)
```

```{r dataAcquisition, include=FALSE, warning = FALSE, tidy = TRUE}
###DATA ACQUISITION
#Below code will import the data present in CSV format into R and store it in a Data Frame.

Bank<- read.csv2("data/bank-additional-full.csv")
#To view first n lines of Bank Variable
head(Bank)
summary(Bank)
#Checking null values in Bank data set
sum(is.na(Bank))
#To display the internal structure of Bank
str(Bank)

```

```{r dataCleaning, include=FALSE, warning = FALSE, tidy = TRUE}
##Data Cleaning
#converting continous data into approprate data type
Bank$emp.var.rate <- as.numeric(Bank$emp.var.rate)
Bank$cons.price.idx <- as.numeric(Bank$cons.price.idx)
Bank$cons.conf.idx <- as.numeric(Bank$cons.conf.idx)
Bank$nr.employed <- as.numeric(Bank$nr.employed)
Bank$euribor3m <- as.numeric(Bank$euribor3m)

#The changes happend above can be seen
summary(Bank)
str(Bank)
```



##Correlation Matrix
Positive correlations are displayed in blue and negative correlations in red color. <br>
Color intensity is proportional to the correlation coefficients.<br>
It is very useful to highlight the most correlated variables in a data table.<br>
In this plot, correlation coefficients is colored according to the value.
```{r correlationMatrix, echo= FALSE }
Bank$y <- as.numeric(Bank$y)
ContinousData<- data.frame(Bank$age, Bank$duration,Bank$campaign,Bank$pdays
               ,Bank$previous,Bank$emp.var.rate,Bank$cons.price.idx
               ,Bank$cons.conf.idx,Bank$euribor3m, Bank$y)
Dependency <- cor(ContinousData)
corrplot(Dependency, method = "color") 
```

From the above correlation matrix, it shows a lot of useful information: <br>
**Example:** the client subscribed a term deposit, is highly negatively correleted to **pdays** and **euribor3m**. <br>
By looking at the colors we can easily know how continous variables are behaving with each other.

```{r warning=FALSE, include=FALSE, tidy=FALSE}
#Re-factoring y
Bank$y <- factor(Bank$y, labels=c("no", "yes"))
#Checking the levels
levels(Bank$y)
str(Bank)

################################# END ###################################

#assigning background color and border line
fill <- "gold1"
line <- "goldenrod2"
```

## Age Distribution
```{r echo=FALSE, fig.height=10, fig.width=10, warning=FALSE, tidy=FALSE}
#assigning background color and border line
fill <- "gold1"
line <- "goldenrod2"

######################## AGE DISTRIBUTION ###################################

#calling ggplot functions to create a histogram
ggplot(Bank, aes(x=age)) + 
  geom_histogram(fill = fill, colour = line, binwidth = 1) +
  ggtitle("Histogram of Age Distribution") +
  xlab("Age") +
  ylab("Count")

################################ END ######################################
```

Figure above shows that, **Age** is not normally distributed and it lends on the left side and it doesn't show that **Age** have a high effect to the variable **y**. People from all the **Age** category could have subscribed for the term deposit.


## No. of Clients working in Different Department

```{r echo=FALSE, fig.height=10, fig.width=10, warning=FALSE, tidy=FALSE}
############# NO. of Clients working in Different Department##############

#calling ggplot function to create a bar graph
ggplot(data=Bank, aes(x=job)) +
  geom_bar() +
  geom_bar(fill = fill, colour = line) +
  geom_text(stat='count', aes(label=..count..), vjust=-1) +
  ggtitle("Client Employment Category")+
  xlab("Employment") +
  ylab("Count")

################################ END ######################################
```

Figure above shows that, most of the clients are admin, blue collar job, and technician respectively.

## Marital status of clients

```{r echo=FALSE, fig.height=10, fig.width=10, warning=FALSE, tidy=FALSE}
######################### Marital status of clients #######################

#calling ggplot function to create a bar graph
ggplot(data=Bank, aes(x=marital)) +
  geom_bar() +
  geom_bar(fill = fill, colour = line) +
  geom_text(stat='count', aes(label=..count..), vjust=-1) +
  ggtitle("Client Marital Status") +
  xlab("Marital Status") +
  ylab("Count")

################################ END ######################################
```

Figure above shows that, most of our client are married. Having the possibility of family members to look after. 

##Education Distribution of clients
```{r echo=FALSE, fig.height=10, fig.width=10, warning=FALSE, tidy=FALSE}
####################### Education Distribution of clients ################

#calling ggplot function to create a bar graph
ggplot(data=Bank, aes(x=education)) +
  geom_bar() +
  geom_bar(fill = fill, colour = line) +
  geom_text(stat='count', aes(label=..count..), vjust=-1) +
  ggtitle("Client Educational Background")+
  xlab("Educational Category") +
  ylab("Count")

################################ END ######################################
```

Figure above shows that, most of the clients having a university degree. Hence, the literacy rate is high amoung the clients. 

## Ways to Contact Client
```{r echo=FALSE, fig.height=10, fig.width=10, warning=FALSE, tidy=FALSE}
############################## Ways to Contact Client #####################

#calling barplot function to create a bar graph
ggplot(data=Bank, aes(x=contact)) +
  geom_bar() +
  geom_bar(fill = fill, colour = line) +
  geom_text(stat='count', aes(label=..count..), vjust=-1) +
  ggtitle("Client Contact Category")+
  xlab("Contact Category") +
  ylab("Count")

################################ END ######################################
```
 
Figure above shows that, there are two ways to contact clients and cellular mode of contact are most preferred.

## Months Count
```{r echo=FALSE, fig.height=10, fig.width=10, warning=FALSE, tidy=FALSE}
############################# Months Count ################################
#calling ggplot function to create a bar graph
ggplot(data=Bank, aes(x=month)) +
  geom_bar() +
  geom_bar(fill = fill, colour = line) +
  geom_text(stat='count', aes(label=..count..), vjust=-1) +
  ggtitle("Month Last Contacted")+
  xlab("Months") +
  ylab("Count")

################################ END ######################################
```

Figure above shows that, top 3 months of most number of client contacted are May, July, and August respectively.

## Education level Vs Getting loans

```{r echo=FALSE, fig.height=10, fig.width=11, warning=FALSE, tidy=FALSE}
###to see people of different education level has loans or dont have loans##
ggplot(data=Bank, aes(x=education,fill = loan)) +
  geom_bar(position=position_dodge()) +
  scale_fill_brewer(palette="Paired") + 
  ggtitle("Education Background Vs Loan") +
  theme_minimal() +
  xlab("Education") +
  ylab("Count")

################################ END ######################################
```

Figure above shows that, the education level of clients who has/hasn't taken the loan. It shows that most clients are having the education level of university degree. 
And it also shows that, most client who has taken loan and the least client who hasn't taken loan are both from the University degree group.

##Impact of previous campain on term deposit

```{r echo=FALSE, fig.height=10, fig.width=10, warning=FALSE, tidy=FALSE}
################# Impact of previous campain on term deposit ##############
ggplot(data=Bank, aes(x=poutcome,fill = y)) +
  geom_bar(position=position_dodge()) +
  scale_fill_brewer(palette="Paired") + 
  ggtitle("Education Background Vs Loan") +
  theme_minimal() +
  xlab("Campain Results") +
  ylab("Count")

################################ END ######################################
```
 
Figure above shows, how the previous campain affects on the term deposit.<br>
It shows thats the last campain was not that successfull and the failure ration is really high.

## Decision Tree Model

```{r dtreeModel, include =FALSE, warning = FALSE, tidy = TRUE}
# Splitting data into training and test datasets. 
set.seed(3)
id <- sample(2,nrow(Bank), prob <- c(0.7,0.3) ,replace = TRUE)
TrainBank <- Bank[id==1,]
TestBank <- Bank[id ==2,]

#No. of rows in each training and testing data set
nrow(TrainBank)
nrow(TestBank)

#Display the internal structure of training and test dataset
str(TrainBank)
str(TestBank)

#Getting column names
colnames(Bank)

#Making our model using train data
Bank_model<- rpart(y~., data = TrainBank)
Bank_model
```

# Decision tree
```{r echo= FALSE, warning = FALSE, tidy = TRUE}
fancyRpartPlot(Bank_model, palettes=c("Greys", "Oranges"))
```

# Prediction Model
```{r include=FALSE, warning = FALSE, tidy = TRUE}
Bank_prediction <- predict(Bank_model, TestBank, type = "class")
Prediction <- table(Bank_prediction,TestBank$y)
```

# Confusion matrix Report
```{r echo=FALSE, warning = FALSE, tidy = TRUE}
confusionMatrix(Prediction)
```
# so our model is more than 90% accurate.
From the matrix, it shows that there are **two possible predicted classes**: "yes" and "no". Predicting whether the client subscribed a term deposit or not, "yes" would mean they have subscribed, and "no" would mean they don't.<br>

Terms:
**true positives (TP):** These are cases in which we predicted yes (they subscribed), and they do subscribed. i.e. **682** <br>
**true negatives (TN):** We predicted no, and they don't subscribed. i.e. **10453**<br>
**false positives (FP):** We predicted yes, but they don't actually subscribed. (Also known as a "Type I error.") i.e. **726**<br>
**false negatives (FN):** We predicted no, but they actually do subscribed. (Also known as a "Type II error.") i.e. **378**<br>

**Accuracy:** It will tell how accurate your model is.<br>
**Sensitivity:** It is also know as True positive rate. <br>
**Specificity:** It is also know as True Negative rate. <br>
**Positive Predictive Value (PPV):** when the prediction is positive, how often is it correct. <br>
**Negative Predictive Value (NPV):** when the prediction is negative, how often is it correct. <br>

#Pruning our decision tree
calculating the optimal prunings based on the cp value.
```{r echo=FALSE, warning = FALSE, tidy = TRUE}
printcp(Bank_model)
```

This function provides the optimal prunings based on the cp value.<br>
The below mentioned list of cp values, we can select the one having the least cross-validated error and use it to prune the tree.<br>
The value of cp should be least, so that the cross-validated error rate is minimum.<br>
Plotcp() provides a graphical representation to the cross validated error summary. The cp values are plotted against the geometric mean to depict the deviation until the minimum value is reached.<br>

```{r echo=FALSE, warning = FALSE, tidy = TRUE}
plotcp(Bank_model)
```

#Prune the tree to create an optimal decision tree :
```{r echo=FALSE, warning = FALSE, tidy = TRUE}
ptree<- prune(Bank_model,
               cp= Bank_model$cptable[which.min(Bank_model$cptable[,"xerror"]),"CP"])
fancyRpartPlot(ptree, uniform=TRUE,
              main="Pruned Tree", palettes=c("Greys", "Oranges"))
```

Note: Not much changes after pruning

#Conclusions:
The results show that the model are fitted to evaluate train data considering that errors is so low and the accuracy in the test set is high.
Decision tress shows that, there will be 81% of client not subscribed, if the nr.employed is >= 6.5 and duration is <608.